Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-11-15 paper/paper.tex
made the useful suggestions from eLife reviews

%%% start IFNrare figure %%%
\begin{figure}
\centerline{
{\bf \Large A}\includegraphics[width=0.28\textwidth,valign=t]{figures/IFN_stochastic/RussellVenn/venn_diagram.pdf}
\hspace{0.02\textwidth}
{\bf \Large B}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large C} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
IFN expression is rare in influenza-infected cells.
{\bf (A)} Number of infected and IFN+ A549 cells in our prior single-cell transcriptomics~\citep{russell2018extreme}.
{\bf (B)} An A549 reporter cell line to identify cells expressing IFN.
The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\FIGDATA[IFNrare]{reporter_sequences}).
The reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\FIGSUPP[IFNrare]{reporter_validation}).
{\bf (C)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the reporter cell line (see \FIGSUPP[IFNrare]{flow} for details).
The plot also shows uninfected cells, and cells infected with saturating amounts of Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}

@@

\caption{
Approach for combined transcriptomic and viral sequencing of single influenza-infected cells that express IFN.
{\bf (A)}
IFN reporter A549 cells are infected with a mix of wild-type and synonymously barcoded viruses.
IFN+ cells are enriched by MACS (\FIGSUPP[workflow]{MACS}), and pooled with non-enriched cells and uninfected canine cells that serve as an internal control for multiplets and mRNA leakage.
{\bf (B)}
The mRNAs from individual cells are converted to cDNAs tagged with cell-specific barcodes.
{\bf (C)}
Cellular transcriptomes are quantified using standard single-cell 3'-end Illumina sequencing, and 
{\bf (D)}
viral genes are enriched by influenza-specific PCR and fully sequenced by PacBio.
{\bf (E)}
The result is a matrix giving the expression of each gene in each cell, as well as the full sequences of the viral genes in infected cells.
}
\label{fig:workflow}

@@

\caption{
Single-cell transcriptomics of IFN-enriched influenza-infected cells.
{\bf (A)} 
Number of cells for which transcriptomes were obtained.
From these numbers, we estimate~\citep{bloom2018estimating} that $\approx$11\% of the transcriptomes are actually multiplets derived from multiple cells. 
{\bf (B)} The number of cellular and viral mRNAs detected for each cell is plotted as a point.
Green lines show the distribution of cellular mRNAs per cell.
Cells outside the dashed magenta lines have unusually low or high amounts of cellular mRNA (likely low-quality emulsions or multiplets), and are excluded from subsequent analyses.
{\bf (C)} Distribution across cells of the fraction of all mRNA derived from influenza.
Cells called as infected are in blue, while other cells are in green.
The inset shows the amount of viral mRNA in the human cells that are called as infected.
{\bf (D)} Number of influenza genes detected per infected cell, and the amount of viral mRNA in cells expressing each number of viral genes.
The majority of cells express all eight gene segments, but a substantial minority fail to express at least one gene.
\FIGSUPP[transcriptomics]{frac_has_gene} shows the frequency that each viral gene is detected.
{\bf (E)} Relative expression of viral genes among infected cells.
{\bf (F)} Number of cells infected with wild-type virus, synonymously barcoded virus, or both.
From the cells infected with both viral barcodes, we estimate~\citep{bloom2018estimating} that 63\% of infected cells are co-infected.
{\bf (G)} Fraction of cellular mRNA from IFN across cells, faceted by whether the cells are infected.
Cells to the left of the first dashed magenta line are classified as IFN-, and cells to the right of the second line are classified as IFN+.
A pseudocount is added to the number of IFN transcripts detected in each cell, which is why none of the fractions are zero.
Many cells that do not express IFN still express ISGs (\FIGSUPP[transcriptomics]{ISG}).
}
\label{fig:transcriptomics}

@@

\centerline{\includegraphics[width=0.85\textwidth]{figures/Validation_Figure/ifn_plot.pdf}}
\caption{Validation that IFN induction is increased by some of the mutations identified in the single-cell virus sequencing of IFN+ cells.
{\bf (A)}
Percent of cells that become IFN+ after infection with a bulk stock of the indicated viral mutant, as determined using a reporter cell line (see \FIGSUPP[validation]{SNPflow} for details).
The numbers indicate the median of four measurements for each viral mutant.
The limit of detection of 0.05\% is indicated with a dashed green line, and the median value for wild type is indicated with a dashed blue line.
Points are colored orange if the mutant virus stock induces IFN more frequently than the wild-type viral stock (one-sided t-test, $P < 0.01$), and blue otherwise.
{\bf (B)}
Similar to the first panel, but validates the increased IFN induction for a large internal deletion in the PB1 gene.
See \FIGSUPP[validation]{delflow} and \FIGSUPP[validation]{delqPCR} for more detailed experimental data.
The experiments in the two panels were performed on different days, and so numerical values can be reliably compared within panels but not between panels.
}
\label{fig:validation}

>>>>

%%% start IFNrare figure %%%
\begin{figure}
\centerline{
{\bf \Large A}\includegraphics[width=0.28\textwidth,valign=t]{figures/IFN_stochastic/RussellVenn/venn_diagram.pdf}
\hspace{0.02\textwidth}
{\bf \Large B}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large C} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
IFN expression is rare in influenza-infected cells.
{\bf (A)} Number of infected and IFN+ A549 cells in our prior single-cell transcriptomics~\citep{russell2018extreme}.
{\bf (B)} An A549 reporter cell line to identify cells expressing IFN.
The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\FIGDATA[IFNrare]{reporter_sequences}).
The reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\FIGSUPP[IFNrare]{reporter_validation}).
{\bf (C)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the type III IFN reporter cell line (see \FIGSUPP[IFNrare]{flow} for details).
The plot also shows uninfected cells, and cells infected with saturating amounts of Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}

@@

\caption{
Approach for combined transcriptomic and viral sequencing of single influenza-infected cells that express IFN.
{\bf (A)}
IFN reporter A549 cells are infected with a mix of wild-type and synonymously barcoded viruses.
IFN+ cells are enriched by MACS (\FIGSUPP[workflow]{MACS}), and pooled with non-enriched cells and uninfected canine cells that serve as an internal control for multiplets and mRNA leakage.
{\bf (B)}
The mRNAs from individual cells are converted to cDNAs tagged with cell-specific barcodes.
{\bf (C)}
Cellular transcriptomes are quantified using standard single-cell 3'-end Illumina sequencing, and 
{\bf (D)}
viral genes are enriched by influenza-specific PCR and fully sequenced by PacBio (in this schematic, only the cell labeled by the red barcode is infected and has viral transcripts that are enriched and sequenced by PacBio).
{\bf (E)}
The result is a matrix giving the expression of each gene in each cell, as well as the full sequences of the viral genes in infected cells.
}
\label{fig:workflow}

@@

\caption{
Single-cell transcriptomics of IFN-enriched influenza-infected cells.
{\bf (A)} 
Number of cells for which transcriptomes were obtained.
From these numbers, we estimate~\citep{bloom2018estimating} that $\approx$11\% of the transcriptomes are actually multiplets derived from multiple cells. 
{\bf (B)} The number of cellular and viral mRNAs detected for each cell is plotted as a point.
Green lines show the distribution of cellular mRNAs per cell.
Cells outside the dashed magenta lines have unusually low or high amounts of cellular mRNA (likely low-quality emulsions or multiplets), and are excluded from subsequent analyses.
{\bf (C)} Distribution across cells of the fraction of all mRNA derived from influenza.
Cells called as infected are in blue, while other cells are in green.
The inset shows the amount of viral mRNA in the human cells that are called as infected.
{\bf (D)} Number of influenza genes detected per infected cell, and the amount of viral mRNA in cells expressing each number of viral genes.
The majority of cells express all eight gene segments, but a substantial minority fail to express at least one gene.
\FIGSUPP[transcriptomics]{frac_has_gene} shows the frequency that each viral gene is detected.
{\bf (E)} Relative expression of viral genes, quantified as the fraction of all viral mRNA in each infected cell that is derived from each gene.
{\bf (F)} Number of cells infected with wild-type virus, synonymously barcoded virus, or both.
From the cells infected with both viral barcodes, we estimate~\citep{bloom2018estimating} that 63\% of infected cells are co-infected.
{\bf (G)} Fraction of cellular mRNA from IFN across cells, faceted by whether the cells are infected.
Cells to the left of the first dashed magenta line are classified as IFN-, and cells to the right of the second line are classified as IFN+.
A pseudocount is added to the number of IFN transcripts detected in each cell, which is why none of the fractions are zero.
Many cells that do not express IFN still express ISGs (\FIGSUPP[transcriptomics]{ISG}).
}
\label{fig:transcriptomics}

@@

\centerline{\includegraphics[width=0.85\textwidth]{figures/Validation_Figure/ifn_plot.pdf}}
\caption{Validation that IFN induction is increased by some of the mutations identified in the single-cell virus sequencing of IFN+ cells.
{\bf (A)}
Percent of infected cells that become IFN+ after infection with a bulk stock of the indicated viral mutant, as determined using a reporter cell line (see \FIGSUPP[validation]{SNPflow} for details).
The numbers indicate the median of four measurements for each viral mutant.
The limit of detection of 0.05\% is indicated with a dashed green line, and the median value for wild type is indicated with a dashed blue line.
Points are colored orange if the mutant virus stock induces IFN more frequently than the wild-type viral stock (one-sided t-test, $P < 0.01$), and blue otherwise.
{\bf (B)}
Similar to the first panel, but validates the increased IFN induction for a large internal deletion in the PB1 gene.
See \FIGSUPP[validation]{delflow} and \FIGSUPP[validation]{delqPCR} for more detailed experimental data.
The experiments in the two panels were performed on different days, and so numerical values can be reliably compared within panels but not between panels.
}
\label{fig:validation}
